---
output: html_document
---
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

Following code computes an average of steps per day.

```{r}
library(plyr)
per_day <- ddply(activity, .(date), function(df) {
  data.frame(date=df$date[1], steps=sum(df$steps, na.rm=T))  
})

step_mean <- mean(per_day$steps, na.rm=T)
```

It shows the mean is `r step_mean`.

## What is the average daily activity pattern?

The following chart shows the change of steps over time.

```{r}
per_step <- ddply(activity, .(interval), function(df) {
  data.frame(interval=df$interval[1], steps=mean(df$steps, na.rm=T))
})
plot(per_step$interval, per_step$steps, type="l", xlab="Interval", ylab="Averate Steps")
```


```{r}
max_step <- max(per_step$steps, na.rm=T)
max_step_interval <- per_step$interval[per_step$steps == max_step]
max_step_interval
```

Also, a person takes the largest number of steps at interval `r max_step_interval` in average.

## Imputing missing values

```{r}
na_counts <- length(which(is.na(activity$step)))
na_counts
```

According to the code above, there are `r na_counts` NA instances. Let's see whether it is reasonable to just drop
these values.

Following code fills the NA records with the average step count.

```{r}
step_mean <- mean(per_step$steps)
filled_activity = activity
filled_activity[is.na(activity$steps),]$steps <- step_mean
filled_per_day <- ddply(filled_activity, .(date), function(df) {
  data.frame(date=df$date[1], steps=sum(df$steps))  
})

filled_steps_mean <- mean(filled_per_day$steps)
filled_steps_median <- median(filled_per_day$steps)
hist(filled_per_day$steps, main="Activity distribution, replacing NA with average", xlab="Step count")
```

Compare the histgram with the one from the table that drops NAs.

```{r}
steps_mean <- mean(per_day$steps)
steps_median <- median(per_day$steps)
hist(per_day$steps, main="Activity distribution, dropping NA", xlab="Step count")
```

The filled version has less portion of smaller and larger number of entries.
So filling with the mean may deemphasize the characteristics at the skirt of the histogram.

Here is the mean and the median of both filled nad dropped versions of tables:

 * Mean (Filling NA with average): `r filled_steps_mean`
 * Median (Filling NA with average): `r filled_steps_median`
 * Mean (Dropping NA): `r steps_mean`
 * Median (Dropping NA): `r steps_median`

They differ only slightly.
That means the distribution isn't skewed in term of the step count.
That corresponds what the histogram shows.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lubridate)

activity$ymd <- ymd(as.character(activity$date))
activity$weekends <- weekdays(activity$ymd) %in% c("Saturday", "Sunday")
per_step_per_daytype <- ddply(activity, .(weekends), function(df) {
  ddply(df, .(interval), function(dg) {
    data.frame(weekends=dg$weekends[1], interval=dg$interval[1], steps=mean(dg$steps, na.rm=T))
  })
})

activity_weekdays <- per_step_per_daytype[!per_step_per_daytype$weekends,]
activity_weekends <- per_step_per_daytype[ per_step_per_daytype$weekends,]
par(mfrow=c(1,2))
plot(activity_weekdays$interval, activity_weekdays$steps, type="l",
     xlab="Interval", ylab="Steps", main="Weekday Activity")
plot(activity_weekends$interval, activity_weekends$steps, type="l",
     xlab="Interval", ylab="Steps", main="Weekend Activity")
```

The chart shows the weekday activity is biased to morning. 
In contrast, the weekend activity is scattered across the day.
